---
title: "simulations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Simulated data

We can simulate data using the **simulate_crispr_screen_data** function. Here, we simulate a dataset with the following parameters:

* *Number of cells*: 1000
* *Probability of gRNA perturbation (absent covariates)*: 0.2
* *Covariates*: log total UMI count (across genes), and percent mitochondrial UMIs. The former covariate is distributed negative-binomially with mean 5,000 and dispersion parameter 2, and the latter is distributed uniformly on [0,2].
* *mRNA mean expression (absent covariates)*: 40
* *gRNA effect on expression*: -4
* *covariate effects*: 1 (for log total UMI count) and 0.5 (for percent mitochondrial UMIs)
* *zero inflation rate*: 0.0
* *negative binomial size parameter*: 2

The covariates and the gRNA presence indicator are dependent; in particular, the gRNA presence indicator is modeled as a logistic GLM of the covariates. The effect of the perturbation on expression is non-zero, meaning that the null hypothesis is false in this example.

```{r, eval=FALSE}
simulated_data <- simulate_crispr_screen_data(num_cells = 1000, 
grna_mean_prob = 0.2, 
covariate_sampler = list(total_umis = function(x) log(rnegbin(x, 1000, 2.5)), 
                         p_mito = function(x) runif(x, 0, 1)),
mRNA_mean_expression = 40, 
gRNA_effect = -4, 
covariate_effects = c(-1, 0.1), 
zero_inflation = 0.0, 
neg_binom_size = 2,
seed = 1234)
```

The output of *simulate_crispr_screen_data* is a list of three elements: the expression vector *Y*, the covariate data frame, and the gRNA presence indicator *X*. Let's peek at these elements.

```{r, eval=FALSE}
map(simulated_data, head)
```

We can run the **sceptre** algorithm on these data to test if we correctly reject the null hypothesis.

```{r, eval=FALSE}
p_value <- run_sceptre_gRNA_gene_pair(expressions = simulated_data$Y, 
                                      gRNA_indicators = simulated_data$X, 
                                      covariate_matrix = simulated_data$covariate_df, 
                                      seed = 1234)
p_value
```

